---
title: ""
author:
- familyname: 
  othernames: 
  address: Monash University
  email: 
  correspondingauthor: true
  qualifications:  

department: Faculty of Business and Statistic
organization: Monash University ETC5513 
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---
```{r,echo = FALSE,message = FALSE,warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(knitr)
library(plotly)
library(gridExtra)
```

**Q1: Which two countries had the highest smoking death rates in 2019？**
**Q2: What are the trends in these two countries over the past decade?**

```{r read data}
death_rate <- read.csv("Data/death-rate-smoking.csv")
```

**(1).The table of top two countries**

```{r toptwotable,echo = FALSE,message = FALSE,warning = FALSE}
rate_rank <- death_rate %>%
  filter(!is.na(Code), Year == 2019) %>%
  rename("smoker_quantity" = "Deaths...Cause..All.causes...Risk..Smoking...Sex..Both...Age..Age.standardized..Rate.") %>%
  rename("Country" = "Entity") %>%
  group_by(Country) %>% 
  summarise(rate = smoker_quantity/100000*100) %>%
  arrange(desc(rate))
top_two <- rate_rank %>%
  head(2) %>%
  kable(caption = "The two countries with the highest death rates from smoking")
top_two
  
```

**(2).The figure of trend of death rate in Kiribati and Solomon Islands**
```{r Kiribati and Solomonfigure,fig.cap = "Death rate in Kiribati and Solomon",echo = FALSE,message = FALSE,warning = FALSE}
Kiribati_rate <- death_rate %>% 
    select(-2) %>% 
    rename("smoker_quantity" =    "Deaths...Cause..All.causes...Risk..Smoking...Sex..Both...Age..Age.standardized..Rate.") %>%
    rename("Country" = "Entity") %>%
    filter(Country == "Kiribati",
           Year >= 2009)
Kiribati_plot <- Kiribati_rate %>%
  ggplot(mapping = aes(x = Year, 
                       y = smoker_quantity)) +
  geom_line() +
  theme_bw() +
  ggtitle("Kiribati trend of death rates from smoking")